---
title: "Bioinformatics Class 11"
author: "Tianhan Xu"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB statistics

Import out PDB statistics CSV file and calculate precent structure by experimental method
```{r}
p <- read.csv("Data Export Summary.csv", row.names=1)
```

```{r}
percent <- (p$Total / sum(p$Total)) *100
names(percent) <- row.names(p)
percent
```


Q1: determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein?

Q2: Type HIV in the search box on the home page and determine how many HIV-1 protease
structures are in the current PDB?

```{r}
library(bio3d)
```

Readf in our HIV-Pr structure
```{r}
pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
str(pdb$atom)
```

```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```
```{r}
# Note that individual $atom records can also be accessed like this  
pdb$atom$elety[1:2]
```
```{r}
# Which allows us to do the following  
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab='B-factor')
```




```{r}
# Print a summary of the coordinate data in $xyz  
pdb$xyz
```

```{r}
# Examine the row and column dimensions  
dim(pdb$xyz)
```

```{r}
pdb$xyz[1, 1:2]
```

Q8. Use the Bio3D   function to write out a protein only PDB file for viewing in VMD. Also write out a second separate PDB file for the ligand with residue name MK1
HINT: In Bio3D you can use the trim.pdb() function together with the pub object and your atom selections such as atom.select(pdb, ’protein’) for the fist file and atom.select(pdb, ‘ligand’) for the second.



```{r}
# Select all C-alpha atoms (return their indices)  
inds.ligand <- atom.select(pdb, "ligand") 
inds.protein <-  atom.select(pdb, 'protein')
inds.protein
```

```{r}
pdb$atom[inds.ligand$atom, ]
```

```{r}
pdb$atom[inds.protein$atom, ]
```

chech we have what we want
```{r}
pdb.ligand <- trim.pdb(pdb, inds= inds.ligand)
pdb.ligand
```


```{r}
write.pdb(pdb.ligand, file= '1hsg_ligand.pdb')
```

```{r}
pdb.protein <- trim.pdb(pdb, inds = inds.protein)
write.pdb(pdb.protein, file = '1hsg_protein.pdb')
```

```{r}
# Download some example PDB files 
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")  
files <- get.pdb(ids, split = TRUE)
```

```{r}
# Extract and align the chains we are interested in  
pdbs <- pdbaln(files, fit = TRUE)
pdbs

```


Q8: What effect does setting the fit=TRUE option have in the related rmsd() function? What does RMSD measure and what would the results indicate if you set fit=FALSE or removed this option? HINT: Bio3D functions have various default options that will be used if the option is not explicitly specified by the user, see help(rmsd) for an example and note that the input options with an equals sign (e.g. fit=FALSE) have default values.

```{r}
# Access the first 5 rows, and 8 columns  
pdbs$ali[1:5, 1:8]
```

```{r}
# Associated residues numbers  
pdbs$resno[1:5, 1:8]
```

```{r}
# Calculate sequence identity 
seqidentity(pdbs)
```


```{r}
# Calculate RMSD  
rd <- rmsd(pdbs)   
# Clustering  
hc <- hclust(as.dist(rd)) 
grps <- cutree(hc, k=3) 
 
# Plot results as dendrogram  
hclustplot(hc, k=3)
```



Section 5: Example Application on Adenylate Kinase (Adk)
```{r}
aa <- get.seq("1ake_A")
```

```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
hits <- plot(b, typ='b')
```
```{r}
hits
```


```{r}
head(hits$pdb.id)
```

```{r}
# Fetch PDBs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
# Align structures
pdbs <- pdbaln(files)
# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)
# Draw schematic alignment
plot(pdbs, labels=ids)
```
Figure 12: Schematic representation of alignment. Grey regions depict aligned residues, while white depict gap regions. The red bar at the top depict sequence conservation.

section 5.4 Principal component analysis
```{r}
# find invariant core
core <- core.find(pdbs)  
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core) 
# Perform PCA
pc.xray <- pca(pdbs)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)  
# Structure-based clustering
hc.rd <- hclust(dist(rd)) 
grps.rd <- cutree(hc.rd, k=3)
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
pc1 <-  mktrj(pc.xray, pc=1, file="pc_1.pdb")
```



