---
title: "Class 9"
author: "Tianhan Xu"
date: "5/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let us read our data
```{r}
url <- "https://bioboot.github.io/bggn213_S18/class-material/WisconsinCancer.csv"
# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(url)
head(wisc.df)
```

```{r}
#how many dianosis are cancer vs non cancer
table(wisc.df$diagnosis)
```
Let's make a new data of matrix in the interest in it
```{r}
wisc.data <- as.matrix(wisc.df[, -c(1,2, 33)])
# Set the row names of wisc.data
rownames(wisc.data) <- wisc.df$id
#x <- c("mary", " barry", "peter")
#x[-1] # which means everything except the first one
#x[-c(1,4)] which means deduct first and forth
```

```{r}
# Create diagnosis vector by completing the missing code
diagnosis <- as.numeric( wisc.df$diagnosis == "M")
sum(diagnosis) # always double check whether you got the right number

```

```{r}
# Q1. How many observations are in this dataset?
dim(wisc.data)
ncol(wisc.data)
# Q2. How many variables/features in the data are suffixed with _mean?
grep("_mean", colnames(wisc.data), value = TRUE, invert = TRUE) # value true give you the name in your data and invert true give you the data that doesn't match the pattern
length(grep("_mean", colnames(wisc.data), value = TRUE, invert = TRUE) )
# Q3. How many of the observations have a malignant diagnosis?
table(wisc.df$diagnosis)
```


```{r}
# Check column means and standard deviations
# if it's hard to look at the table directly, plot it!
plot(colMeans(wisc.data), type='o')

plot(apply(wisc.data,2,sd), type = 'o')

```

```{r}
# you have to ensure there's no NA in your data, otherwise sscal cannot be applied
wisc.pr <- prcomp(wisc.data, scale. = TRUE)

summary(wisc.pr)

```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
A: 0.4427

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
A: 3

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
A:7

```{r}
biplot(wisc.pr)
```

```{r}
# Scatter plot observations by components 1 and 2
plot( wisc.pr$x[, c(1,2)], col=diagnosis + 2, xlab = "PC1", ylab = "PC2")
```

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[, c(1, 3)], col = (diagnosis + 4), 
     xlab = "PC1", ylab = "PC3")
```


## Screen plots showing the proportion of variance explained as the number of principal components increases
```{r}
#Calculate the variance of each principal component by squaring the sdev component of wisc.pr (i.e. wisc.pr$sdev^2). Save the result as an object called pr.var.
pr.var <-  wisc.pr$sdev^2
pve <- pr.var / sum(pr.var)
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")

```


```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), , las=3, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

```{r}
# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```



Section 3.
Hierarchical clustering of case data

```{r}
#scale the data
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist)
plot(wisc.hclust)
abline(h=18, col='red', lwd=3 )
```

```{r}
#selecting numbers of clusters
wisc.hclust.clusters <- cutree(wisc.hclust, h=18)
```

##how do these groups match our `diagnosis`


```{r}
table(wisc.hclust.clusters, diagnosis)

```





Section 4.
K-means clustering and comparing results
```{r}
wisc.km <- kmeans(wisc.data, centers= 5, nstart= 15)
table(wisc.km$cluster, diagnosis)

```

```{r}
table(wisc.km$cluster, wisc.hclust.clusters)
```

Section 5.
Clustering on PCA results
```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
scl <- wisc.pr$x[,1:7]
di <- dist(scl)
wisc.pr.hclust <- hclust(di, method='ward.D2')
plot(wisc.pr.hclust)
```

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
plot(wisc.pr$x[,1:7], col= wisc.pr.hclust.clusters)

```

```{r}
#compare different data you created
table(wisc.pr.hclust.clusters, diagnosis)
table(wisc.km$cluster, wisc.pr.hclust.clusters)
```

```{r}
#install packages("rgl")
library(rgl)
plot3d(wisc.pr$x[,1:7], col=wisc.hclust.clusters)
```

```{r}
url <-  "https://tinyurl.com/new-samples-CSV"
new <-  read.csv(url)
npc <- predict(wisc.pr, newdata = new)
```

```{r}
plot(wisc.pr$x[,1:2], col=wisc.hclust.clusters)
points(npc[1], npc[2], col=c ('purple','blue'), pch=16, cex=3)
```






